---
title: "RNG"
author: "Yaqiong Yao"
date: "10/19/2018"
output: pdf_document
header-includes:
  \usepackage{algorithm}
---

# 5.2.1 Rejection sampling

## 1

\begin{align*}
\int_0^{\infty}(2x^{\theta-1}+x^{\theta-\frac{1}{2}})e^{-x}dx &= 2\int_0^{\infty}x^{\theta-1}e^{-x}dx + \int_0^{\infty}x^{\theta-\frac{1}{2}}e^{-x}dx \\
&= 2\Gamma(\theta)\int_0^{\infty} \frac{1}{\Gamma(\theta)}x^{\theta-1}e^{-x}dx + \Gamma(\theta+\frac{1}{2}) \int_0^{\infty} \frac{1}{\Gamma(\theta+\frac{1}{2})}x^{\theta-\frac{1}{2}}e^{-x}dx \\
&= 2\Gamma(\theta) + \Gamma(\theta+\frac{1}{2})
\end{align*}

Thus $C = \left(2\Gamma(\theta) + \Gamma(\theta+\frac{1}{2})\right)^{-1}$. The components are two gamma distribution $G(\theta, 1)$ and $G(\theta+\frac{1}{2}, 1)$ with weights $2\Gamma(\theta)C$ and $\Gamma(\theta + \frac{1}{2})C$.

## 2

\begin{algorithm}
\caption{Sample from g(x)}
\begin{enumerate}
\item Sample $u \sim \mathcal{U}(0, 1)$
\item Calculate $$ X = I(u < 2\Gamma(\theta)C)Gamma(\theta, 1) + I(u > 2\Gamma(\theta)C)Gamma(\theta + \frac{1}{2}, 1)$$
\end{enumerate}
\end{algorithm}


```{r}
g.sample <- function(n, theta){
  C <- 1/(2*gamma(theta)+gamma(theta + 1/2))
  ind <- runif(n, 0, 1)
  g.sample <- as.numeric(ind < 2*gamma(theta)*C) * rgamma(n, shape = theta, scale = 1) + 
    (1 - as.numeric(ind < 2*gamma(theta)*C)) * rgamma(n, shape = (theta + 1/2), scale = 1)
  return(g.sample)
}
n <- 10000
theta <- 2
set.seed(123)
g.s <- g.sample(n, theta = theta)


gx <- function(x){
  C <- 1/(2*gamma(theta)+gamma(theta + 1/2))
  2*gamma(theta)*C*dgamma(x, shape = theta, scale = 1) + 
    gamma(theta + 1/2)*C*dgamma(x, shape = (theta + 1/2), scale = 1)
}

plot(density(g.s), main = "Kernel density of g and true density", col = "blue", lty = 1)
plot(gx, 0, 12, add = TRUE, col = "red", lty = 5)
legend("topright", col = c("blue", "red"), c("kernel density estimation of g", "true density"),
       lty = c(2, 5))
```

Here, I set $\theta = 2$.

## 3

First, we need to determine a $\alpha$ that satisfies $\alpha = \sup\frac{q(x)}{g(x)}$.

\begin{align*}
\alpha &= \sup\frac{q(x)}{g(x)} \\
&= \sup\frac{\sqrt{4+x}}{C(2+x^{\frac{1}{2}})}
\end{align*}
where $q(x) = \sqrt{4+x}x^{\theta-1}e^{-x}$.

Since $f(x) = \sqrt x$ is a convace function, $\alpha = \sup\frac{q(x)}{g(x)} = \frac{1}{C}$. So, clearly, if we substitue $g(x)$ with $h(x) = (2x^{\theta-1}+x^{\theta-\frac{1}{2}})e^{-x}$. Then we should use $\beta = 1$ instead of $\alpha$.


\begin{algorithm}
\caption{Sample f(x) from g(x)}
\begin{enumerate}
\item Sample $u \sim \mathcal{U}(0, 1)$
\item Sample $X$ from $g(x)$
\item Calculate $$rate = \frac{\sqrt{4 + X}}{2 + X^{\frac{1}{2}}}$$
\item Keep samples let $rate > u$
\end{enumerate}
\end{algorithm}



```{r}
f.rs <- function(n, theta){
  u <- runif(n, 0, 1)
  g_s <- g.sample(n, theta)
  rate <- sqrt(4 + g_s)/(2 + g_s^(1/2))
  f.s <- g_s[u < rate]
  return(f.s)
}

n <- 15000
theta <- 2
f.s <- f.rs(n, theta)[1:10000]


fx <- function(x){
  qx <- function(x) sqrt(4 + x) * (x^(theta-1)) * exp(-x)
  G <- integrate(qx, 0, Inf)
  (sqrt(4 + x) * (x^(theta-1)) * exp(-x))/G$value
}

plot(density(f.s), main = "Kernel density of f and true density", col = "blue", lty = 1)
plot(fx, 0, 12, add = TRUE, col = "red", lty = 5)
legend("topright", col = c("blue", "red"), c("estimated kernel density of f", "true density"),
       lty = c(2, 5))
```


# 5.2.2 Mixture proposal

## 1

Use the mixture beta distribution of $\frac{1}{3}I(\theta\le\beta)Beta(\theta, 1) + \frac{2}{3}I(\theta\le\beta)Beta(1, \theta) + \frac{1}{3}I(\theta>\beta)Beta(\beta, 1) + \frac{2}{3}I(\theta>\beta)Beta(1, \beta)$.

Since 

$$\frac{x^{\theta-1}}{1+x^2} + \sqrt{2 + x^2}(1-x)^{\beta-1} \le x^a + 2(1-x)^{a-1}$$
where $a = min(\theta, \beta)$.

\begin{algorithm}
\caption{Sample f(x) from mixture beta distribution}
\begin{enumerate}
\item Sample $u \sim \mathcal{U}(0, 1)$
\item Set $a = min(\theta, \beta)$
\item Sample $X$ from mixture beta distribution
\item Calculate rate $$\frac{\frac{X^{\theta-1}}{1+X^2} + \sqrt{2 + X^2}(1-X)^{\beta-1}}{X^a + 2(1-X)^{a-1}}$$
\item Keep samples let rate > $u$
\end{enumerate}
\end{algorithm}



```{r}
mix.beta <- function(n, beta, theta){
  a <- min(theta, beta)
  u <- runif(n, 0, 1)
  s <- as.numeric(u > 1/3)*rbeta(n, shape1 = 1, shape2 = a) + 
    (1-as.numeric(u > 1/3))*rbeta(n, shape1 = a, shape2 = 1)
  return(s)
}

mix.rs <- function(n, beta, theta){
  u <- runif(n, 0, 1)
  a <- min(beta, theta)
  mbs <- mix.beta(n, beta, theta)
  rate <- (mbs^(theta-1)/(1+mbs^2) + sqrt(2 + mbs^2)*(1-mbs)^(beta-1))/(mbs^a+2*(1-mbs)^(a-1))
  m.sample <- mbs[u < rate]
  return(m.sample)
}

theta <- 10
beta <- 10
n <- 22000
set.seed(123)
sp <- mix.rs(n, beta, theta)[1:10000]

fx <- function(x){
  qx <- function(x) x^(theta-1)/(1+x^2) + sqrt(2 + x^2)*(1-x)^(beta-1)
  (x^(theta-1)/(1+x^2) + sqrt(2 + x^2)*(1-x)^(beta-1))/integrate(qx, 0, 1)$value
}

plot(density(sp), xlim = c(0,1), main = "Kernel density of f and true density", col = "blue", lty = 1)
plot(fx, 0, 1, add = TRUE, col = "red", lty = 5)
legend("topright", col = c("blue", "red"), c("estimated kernel density of f", "true density"),
       lty = c(2, 5))

```



## 2

Since

$$\frac{x^{\theta-1}}{1+x^2} \le x^{\theta-1}$$
$$\sqrt{2 + x^2}(1-x)^{\beta-1} \le \sqrt{3}(1-x)^{\beta-1}$$

Thus, I use two beta distribution $Beta(\theta, 1)$ and $Beta(1, \beta)$. Here $\alpha_1 = \frac{1}{\theta}$ and $\alpha_2 = \frac{\sqrt{3}}{\beta}$.

\begin{algorithm}
\caption{Sample f(x) seperately}
\begin{enumerate}
\item Sample $u_1 \sim \mathcal{U}(0, 1)$
\item If $u_1 > \frac{\frac{1}{\theta}}{\frac{1}{\theta} + \frac{\sqrt{3}}{\beta}}$, sample $X$ from $Beta(1, \beta)$ and calculate rate using $\frac{1}{1+X^2}$; otherwise, sample $X$ from $Beta(\theta, 1)$ and calculate rate using $\sqrt{\frac{2 + X^2}{3}}$
\item Sample $u_2 \sim \mathcal{U}(0, 1)$
\item Keep samples let rate > $u_2$
\end{enumerate}
\end{algorithm}

```{r}
sep.beta <- function(n, beta, theta){
  X <- rep(NA, n)
  rate <- rep(NA, n)
  u1 <- runif(n, 0, 1)
  ind <- 1/theta + sqrt(3)/beta
  I1 <- u1 > (1/theta)/ind; n1 <- sum(u1 > (1/theta)/ind)
  I2 <- u1 <= (1/theta)/ind; n2 <- n-n1
  X[I1] <- rbeta(n1, shape1 = 1, shape2 = beta)
  X[I2] <- rbeta(n2, shape1 = theta, shape2 = 1)
  
  u2 <- runif(n, 0, 1)
  rate[I1] <- 1/(1 + X[I1]^2)
  rate[I2] <- sqrt((2+X[I2]^2)/3)
  
  s <- X[u2 < rate]
  return(s)
}

theta <- 10
beta <- 10
n <- 12000
set.seed(123)
sp <- sep.beta(n, beta, theta)[1:10000]

plot(density(sp), xlim = c(0,1), main = "Kernel density of f and true density", col = "blue", lty = 1)
plot(fx, 0, 1, add = TRUE, col = "red", lty = 5)
legend("topright", col = c("blue", "red"), c("estimated kernel density of f", "true density"),
       lty = c(2, 5))

```





